---
title: "pilot study data"
output:
  html_document: default
  pdf_document: default
date: "2024-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(themis)
library(tidyverse)
library(tidymodels)
library(tidyclust)
library(ggplot2)
```

```{r}
library(readr)

Artworks <- read_csv("Artworks - Sheet1.csv")


#demographics <- read_csv("demographics.csv")

qualtrics_data <- read_csv("raw_cleaned_data.csv")
```





```{r}
library(dplyr)
library(tidyr)

artworks <- Artworks
```


```{r setup}
library(dplyr)


rating_columns <- names(qualtrics_data)[grepl('rating_scores', names(qualtrics_data))]


is_numeric_or_na <- function(x) {
  suppressWarnings(!is.na(as.numeric(as.character(x)))) | is.na(x)
}


filtered_qualtrics_data <- qualtrics_data %>%
  rowwise() %>%
  filter(all(sapply(c_across(all_of(rating_columns)), is_numeric_or_na))) %>%
  ungroup()

str(filtered_qualtrics_data)

filtered_qualtrics_data<- filtered_qualtrics_data %>% 
  mutate(participantID=row_number())
# view(filtered_qualtrics_data)
```


```{r}
library(dplyr)
library(tidyr)
library(stringr)


qualtrics_long <- filtered_qualtrics_data %>%
  pivot_longer(
    cols = contains('rating_scores'),
    names_to = 'Artwork_ID',
    names_pattern = '(\\d+)_rating_scores', 
    values_to = 'Rating'
  ) %>%
  mutate(Artwork_ID = as.numeric(Artwork_ID))  


head(qualtrics_long)
```


```{r}
artworks$ID <- 1:nrow(artworks)

final_dataset <- merge(artworks, qualtrics_long, by = 'Artwork_ID')


```


Averages
```{r}
library(dplyr)
final_dataset$Rating <- as.numeric(as.character(final_dataset$Rating))

sum(is.na(final_dataset$Rating))

average_ratings_by_style <- final_dataset %>%
  filter(!is.na(Rating) & !is.na(Style)) %>%
  group_by(Style) %>%
  summarize(Average_Rating = mean(Rating, na.rm = TRUE))


print(average_ratings_by_style)

```


```{r}
library(ggplot2)
library(dplyr)

average_ratings_by_style <- final_dataset %>%
  filter(!is.na(Rating) & !is.na(Style)) %>%
  group_by(Style) %>%
  summarize(
    Average_Rating = mean(Rating, na.rm = TRUE) - 50,
    SD_Rating = sd(Rating, na.rm = TRUE),
    n = n() 
  )

average_ratings_by_style <- average_ratings_by_style %>%
  arrange(Average_Rating)

ggplot(average_ratings_by_style, aes(x = reorder(Style, Average_Rating), y = Average_Rating)) +
  geom_bar(stat = "identity", fill = "orange") +
  geom_errorbar(aes(ymin = Average_Rating - SD_Rating/sqrt(n), ymax = Average_Rating + SD_Rating/sqrt(n)), width = 0.2) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +  # Add a red dashed horizontal line at y=50
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Average Liking Ratings Across Art Styles/Movements", x = "Art Style", y = "Average Rating")+
  scale_y_continuous(breaks = seq(-50, 50, 10), limits = c(-25, 20))




```


Top 10s 
```{r}
library(dplyr)

average_ratings_by_piece <- final_dataset %>%
  group_by(Title) %>%
  summarize(Average_Rating = mean(Rating, na.rm = TRUE))

top_liked <- average_ratings_by_piece %>%
  arrange(desc(Average_Rating)) %>%
  slice_head(n = 10)


least_liked <- average_ratings_by_piece %>%
  arrange(Average_Rating) %>%
  slice_head(n = 10)


print(top_liked)
print(least_liked)

```




```{r}

std_dev_by_style <- final_dataset %>%
  group_by(Style) %>%
  summarize(SD_Rating = sd(Rating, na.rm = TRUE))


ggplot(std_dev_by_style, aes(x = Style, y = SD_Rating)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Disagreement in Ratings Across Art Styles/Movements")

```



```{r}

std_dev_by_piece <- final_dataset %>%
  group_by(Title) %>%
  summarize(SD_Rating = sd(Rating, na.rm = TRUE))


divisive_pieces <- std_dev_by_piece %>%
  arrange(desc(SD_Rating)) %>%
  top_n(10)


agreed_on_pieces <- std_dev_by_piece %>%
  arrange(SD_Rating) %>%
  top_n(-10)

print(divisive_pieces)
print(agreed_on_pieces)
```



```{r}
library(ggplot2)
library(dplyr)


std_dev_by_piece <- final_dataset %>%
  group_by(Title) %>%
  summarize(SD_Rating = sd(Rating, na.rm = TRUE))


divisive_pieces1 <- std_dev_by_piece %>%
  top_n(6, SD_Rating)


agreed_on_pieces1 <- std_dev_by_piece %>%
  top_n(6, -SD_Rating)


divisive_data <- final_dataset %>%
  filter(Title %in% divisive_pieces1$Title)

agreed_on_data <- final_dataset %>%
  filter(Title %in% agreed_on_pieces1$Title)



ggplot(divisive_data, aes(x = Rating)) +
  geom_histogram(bins = 10, fill="red") +
  facet_wrap(~Title, scales = "free_y", nrow = 3, ncol = 2) +
  labs(title = "Histogram of Ratings for the Most Divisive Pieces")+
  scale_y_continuous(breaks = seq(-50, 50, 10))

ggplot(agreed_on_data, aes(x = Rating)) +
  geom_histogram(bins = 10, fill= "green") +
  facet_wrap(~Title, scales = "free_y", nrow = 3, ncol = 2) +  
  labs(title = "Histogram of Ratings for the Most Agreed Upon Pieces")+
  scale_y_continuous(breaks = seq(-50, 50, 10))


```



```{r}

library(dplyr)
library(ggplot2)

sd_by_title <- final_dataset %>%
  group_by(Title) %>%
  summarise(ratings_sd = sd(Rating, na.rm = TRUE))


clean_dataset <- final_dataset 


ggplot(clean_dataset, aes(x = Rating)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +  
  xlab("Rating Scores") +
  ylab("Count") +
  ggtitle("Histogram of Rating Variability Across Rating Scores")


```
```{r}


Artworks1 <- Artworks %>%
  mutate(ID = row_number())

merged_data <- recomm_data %>%
  left_join(Artworks1, by = "ID") %>%
  select(-ID) %>%
  rename("link to image" = "link to image")


merged_data1<- merged_data%>%
  select(participantID, Rating, "link to image")

pivoted_qualdata <- merged_data1 %>%
  pivot_wider(names_from = participantID, values_from = Rating, names_prefix = "participant_")

print(pivoted_qualdata)
```


## Recommender System


### Data prep
```{r}
# recomm_data<- filtered_qualtrics_data %>%
#   select(-StartDate, -EndDate, -Progress, -`Duration (in seconds)`, -Finished, -RecordedDate, -DistributionChannel, -UserLanguage, -consent, -visits, -art_knowledge, -opens_1, -opens_2, -opens_3, -edu, -age, -gender, -`gender _4_TEXT`, -participantID)

# This should get you started -- DP 2024-04-30
recomm_data <- final_dataset %>% 
  select(participantID, ID, Rating)

recomm_data_wide <- recomm_data %>% 
  spread(ID, Rating) 
```



```{r}

library(recommenderlab)


ratings_data <- as(as.matrix(recomm_data_wide), "realRatingMatrix")

train <- ratings_data[1:75,]
```



```{r}
scheme <- evaluationScheme(train, method = "cross-validation", k = 10, given = -5, goodRating = 50)

algorithms <- list(
  `random items` = list(name = "RANDOM", param = NULL), 
  `popular items` = list(name = "POPULAR", param = NULL),
  `user-based CF` = list(name = "UBCF", param = list(nn = 3)),
  `item-based CF` = list(name = "IBCF", param = list(k = 10)),
  `funk-svd-hd` = list(name ="SVDF", param = list(k = 10)),
  `funk-svd-ld` = list(name ="SVDF", param = list(k = 3))
  )

results <- evaluate(scheme, algorithms, type = "ratings", progress = FALSE)


```

All recommender approaches do better than random, so at least there's that. Funk-svd seems to do somewhat better than "popular"-based recommender systems. So there is some hint of personalizability here. 


```{r}

library(stringr)


get_MAE <- function(res){
  map(res, ~.x[3])
}

res_getter <- function(results){
  n <- names(results)
  reslist <- lapply(results, getResults)
  v <- map(reslist, get_MAE) %>% unlist()
  
  tibble(
    alg = names(v),
    val = v
  ) %>% 
    mutate(
      fold = str_extract(alg, "\\d{1,2}(?!.*\\d{1,2})"),
      alg = str_remove(alg, fold)
      )
    
  # lapply(reslist, ~tibble(.x[1,]))
}           

res_getter(results) %>% 
  ggplot(aes(x=alg, y = val)) +
  geom_boxplot() +
  coord_flip() +
  labs(y = "MAE")

# x <- lapply(results, getResults)
# map2(names(results), results, ~as_tibble(getResults(.y)) %>% mutate(namez = .x))
```

In terms of mean absolute error, funk-svd is on the order of 15-17. On a scale of 100 that's not *too* bad, but it's not super great -- perhaps more data is needed.



### Hyperparameter tuning


```{r}
algorithms_h <- list(
  `funk-svd-ld` = list(name ="SVDF", param = list(k = 3)),  
 `low-lr-high-reg` = list(name = "SVDF", param = list(
    k = 3,
    lr_all = 0.0001,
    reg_all = 0.02,
    n_epochs = 50
  )),
  `medium-lr-medium-reg` = list(name = "SVDF", param = list(
    k = 3,
    lr_all = 0.0003,
    reg_all = 0.015,
    n_epochs = 50
  )),
  `high-lr-low-reg` = list(name = "SVDF", param = list(
    k = 3,
    lr_all = 0.0007,
    reg_all = 0.01,
    n_epochs = 50
  )),
  `low-lr-more-epochs` = list(name = "SVDF", param = list(
    k = 3,
    lr_all = 0.0001,
    reg_all = 0.01,
    n_epochs = 100
  )),
  `high-lr-more-epochs` = list(name = "SVDF", param = list(
    k = 3,
    lr_all = 0.0005,
    reg_all = 0.005,
    n_epochs = 100
  ))
)
results_h <- evaluate(scheme, algorithms_h, type = "ratings", progress = FALSE)

get_MAE <- function(res){
  map(res, ~.x[3])
}

res_getter_h <- function(results_h){
  n <- names(results_h)
  reslist <- lapply(results_h, getResults)
  v <- map(reslist, get_MAE) %>% unlist()
  
  tibble(
    alg = names(v),
    val = v
  ) %>% 
    mutate(
      fold = str_extract(alg, "\\d{1,2}(?!.*\\d{1,2})"),
      alg = str_remove(alg, fold)
      )
    
  # lapply(reslist, ~tibble(.x[1,]))
}           

res_getter(results_h) %>% 
  ggplot(aes(x=alg, y = val)) +
  geom_boxplot() +
  coord_flip() +
  labs(y = "MAE")
```

I tried other tuning options but nothing performed better than the normal one. More extreme learning rates or regularization led to no results and there seems to be no gains from minor changes. 
