---
title: "pilot study data"
output:
  html_document: default
  pdf_document: default
date: "2024-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, include=FALSE}
library(themis)
library(tidyverse)
library(tidymodels)
library(tidyclust)
library(ggplot2)
```

```{r}
library(readr)

Artworks <- read_csv("Artworks - Sheet1.csv")


#demographics <- read_csv("demographics.csv")

qualtrics_data <- read_csv("clean_complete_pilot_data.csv")
```





```{r}
library(dplyr)
library(tidyr)

artworks <- Artworks
```


```{r}
qualtrics_data <- qualtrics_data %>%
  slice(-c(3:8))
view(qualtrics_data)
```


```{r setup}
library(dplyr)


rating_columns <- names(qualtrics_data)[grepl('rating_scores', names(qualtrics_data))]


is_numeric_or_na <- function(x) {
  suppressWarnings(!is.na(as.numeric(as.character(x)))) | is.na(x)
}


filtered_qualtrics_data <- qualtrics_data %>%
  rowwise() %>%
  filter(all(sapply(c_across(all_of(rating_columns)), is_numeric_or_na))) %>%
  ungroup()

str(filtered_qualtrics_data)

filtered_qualtrics_data<- filtered_qualtrics_data %>% 
  mutate(participantID=row_number())
 view(filtered_qualtrics_data)
```


```{r}
library(dplyr)
library(tidyr)
library(stringr)


qualtrics_long <- filtered_qualtrics_data %>%
  pivot_longer(
    cols = contains('rating_scores'),
    names_to = 'Artwork_ID',
    names_pattern = '(\\d+)_rating_scores', 
    values_to = 'Rating'
  ) %>%
  mutate(Artwork_ID = as.numeric(Artwork_ID))  


view(qualtrics_long)
```


```{r}
artworks$ID <- 1:nrow(artworks)

final_dataset <- merge(artworks, qualtrics_long, by = 'Artwork_ID')


```


Averages
```{r}
library(dplyr)
final_dataset$Rating <- as.numeric(as.character(final_dataset$Rating))

sum(is.na(final_dataset$Rating))

average_ratings_by_style <- final_dataset %>%
  filter(!is.na(Rating) & !is.na(Style)) %>%
  group_by(Style) %>%
  summarize(Average_Rating = mean(Rating, na.rm = TRUE))


print(average_ratings_by_style)

```


```{r}
library(ggplot2)
library(dplyr)

average_ratings_by_style <- final_dataset %>%
  filter(!is.na(Rating) & !is.na(Style)) %>%
  group_by(Style) %>%
  summarize(
    Average_Rating = mean(Rating, na.rm = TRUE) - 50,
    SD_Rating = sd(Rating, na.rm = TRUE),
    n = n() 
  )

average_ratings_by_style <- average_ratings_by_style %>%
  arrange(Average_Rating)

ggplot(average_ratings_by_style, aes(x = reorder(Style, Average_Rating), y = Average_Rating)) +
  geom_bar(stat = "identity", fill = "orange") +
  geom_errorbar(aes(ymin = Average_Rating - SD_Rating/sqrt(n), ymax = Average_Rating + SD_Rating/sqrt(n)), width = 0.2) +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +  # Add a red dashed horizontal line at y=50
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Average Liking Ratings Across Art Styles/Movements", x = "Art Style", y = "Average Rating")+
  scale_y_continuous(breaks = seq(-50, 50, 10), limits = c(-25, 20))




```


Top 10s 
```{r}
library(dplyr)

average_ratings_by_piece <- final_dataset %>%
  group_by(Title) %>%
  summarize(Average_Rating = mean(Rating, na.rm = TRUE))

top_liked <- average_ratings_by_piece %>%
  arrange(desc(Average_Rating)) %>%
  slice_head(n = 10)


least_liked <- average_ratings_by_piece %>%
  arrange(Average_Rating) %>%
  slice_head(n = 10)


print(top_liked)
print(least_liked)

```




```{r}

std_dev_by_style <- final_dataset %>%
  group_by(Style) %>%
  summarize(SD_Rating = sd(Rating, na.rm = TRUE))


ggplot(std_dev_by_style, aes(x = Style, y = SD_Rating)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Disagreement in Ratings Across Art Styles/Movements")

```



```{r}

std_dev_by_piece <- final_dataset %>%
  group_by(Title) %>%
  summarize(SD_Rating = sd(Rating, na.rm = TRUE))


divisive_pieces <- std_dev_by_piece %>%
  arrange(desc(SD_Rating)) %>%
  top_n(10)


agreed_on_pieces <- std_dev_by_piece %>%
  arrange(SD_Rating) %>%
  top_n(-10)

print(divisive_pieces)
print(agreed_on_pieces)
```



```{r}
library(ggplot2)
library(dplyr)


std_dev_by_piece <- final_dataset %>%
  group_by(Title) %>%
  summarize(SD_Rating = sd(Rating, na.rm = TRUE))


divisive_pieces1 <- std_dev_by_piece %>%
  top_n(6, SD_Rating)


agreed_on_pieces1 <- std_dev_by_piece %>%
  top_n(6, -SD_Rating)


divisive_data <- final_dataset %>%
  filter(Title %in% divisive_pieces1$Title)

agreed_on_data <- final_dataset %>%
  filter(Title %in% agreed_on_pieces1$Title)



ggplot(divisive_data, aes(x = Rating)) +
  geom_histogram(bins = 10, fill="red") +
  facet_wrap(~Title, scales = "free_y", nrow = 3, ncol = 2) +
  labs(title = "Histogram of Ratings for the Most Divisive Pieces")+
  scale_y_continuous(breaks = seq(-50, 50, 10))

ggplot(agreed_on_data, aes(x = Rating)) +
  geom_histogram(bins = 10, fill= "green") +
  facet_wrap(~Title, scales = "free_y", nrow = 3, ncol = 2) +  
  labs(title = "Histogram of Ratings for the Most Agreed Upon Pieces")+
  scale_y_continuous(breaks = seq(-50, 50, 10))


```



```{r}

library(dplyr)
library(ggplot2)

sd_by_title <- final_dataset %>%
  group_by(Title) %>%
  summarise(ratings_sd = sd(Rating, na.rm = TRUE))


clean_dataset <- final_dataset 


ggplot(clean_dataset, aes(x = Rating)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +  
  xlab("Rating Scores") +
  ylab("Count") +
  ggtitle("Histogram of Rating Variability Across Rating Scores")


```
```{r}


Artworks1 <- Artworks %>%
  mutate(ID = row_number())

merged_data <- recomm_data %>%
  left_join(Artworks1, by = "ID") %>%
  select(-ID) %>%
  rename("link to image" = "link to image")


merged_data1<- merged_data%>%
  select(participantID, Rating, "link to image")

pivoted_qualdata <- merged_data1 %>%
  pivot_wider(names_from = participantID, values_from = Rating, names_prefix = "participant_")

print(pivoted_qualdata)
```


## Recommender System


### Data prep
```{r}
# recomm_data<- filtered_qualtrics_data %>%
#   select(-StartDate, -EndDate, -Progress, -`Duration (in seconds)`, -Finished, -RecordedDate, -DistributionChannel, -UserLanguage, -consent, -visits, -art_knowledge, -opens_1, -opens_2, -opens_3, -edu, -age, -gender, -`gender _4_TEXT`, -participantID)

# This should get you started -- DP 2024-04-30
recomm_data <- final_dataset %>% 
  select(participantID, ID, Rating)

recomm_data_wide <- recomm_data %>% 
  spread(ID, Rating) 
```



```{r}

library(recommenderlab)


ratings_data <- as(as.matrix(recomm_data_wide), "realRatingMatrix")

train <- ratings_data[1:75,]
```



```{r}
scheme <- evaluationScheme(train, method = "cross-validation", k = 10, given = -5, goodRating = 50)

algorithms <- list(
  `random items` = list(name = "RANDOM", param = NULL), 
  `popular items` = list(name = "POPULAR", param = NULL),
  `user-based CF` = list(name = "UBCF", param = list(nn = 3)),
  `item-based CF` = list(name = "IBCF", param = list(k = 10)),
  `funk-svd-hd` = list(name ="SVDF", param = list(k = 10)),
  `funk-svd-ld` = list(name ="SVDF", param = list(k = 3))
  )

results <- evaluate(scheme, algorithms, type = "ratings", progress = FALSE)


```

All recommender approaches do better than random, so at least there's that. Funk-svd seems to do somewhat better than "popular"-based recommender systems. So there is some hint of personalizability here. 


```{r}

library(stringr)


get_MAE <- function(res){
  map(res, ~.x[3])
}

res_getter <- function(results){
  n <- names(results)
  reslist <- lapply(results, getResults)
  v <- map(reslist, get_MAE) %>% unlist()
  
  tibble(
    alg = names(v),
    val = v
  ) %>% 
    mutate(
      fold = str_extract(alg, "\\d{1,2}(?!.*\\d{1,2})"),
      alg = str_remove(alg, fold)
      )
    
  # lapply(reslist, ~tibble(.x[1,]))
}           

res_getter(results) %>% 
  ggplot(aes(x=alg, y = val)) +
  geom_boxplot() +
  coord_flip() +
  labs(y = "MAE")

# x <- lapply(results, getResults)
# map2(names(results), results, ~as_tibble(getResults(.y)) %>% mutate(namez = .x))
```

In terms of mean absolute error, funk-svd is on the order of 15-17. On a scale of 100 that's not *too* bad, but it's not super great -- perhaps more data is needed.



### Hyperparameter tuning


```{r}
algorithms_h <- list(
  `funk-svd-ld` = list(name ="SVDF", param = list(k = 3)),  
 `low-lr-high-reg` = list(name = "SVDF", param = list(
    k = 3,
    lr_all = 0.0001,
    reg_all = 0.02,
    n_epochs = 50
  )),
  `medium-lr-medium-reg` = list(name = "SVDF", param = list(
    k = 3,
    lr_all = 0.0003,
    reg_all = 0.015,
    n_epochs = 50
  )),
  `high-lr-low-reg` = list(name = "SVDF", param = list(
    k = 3,
    lr_all = 0.0007,
    reg_all = 0.01,
    n_epochs = 50
  )),
  `low-lr-more-epochs` = list(name = "SVDF", param = list(
    k = 3,
    lr_all = 0.0001,
    reg_all = 0.01,
    n_epochs = 100
  )),
  `high-lr-more-epochs` = list(name = "SVDF", param = list(
    k = 3,
    lr_all = 0.0005,
    reg_all = 0.005,
    n_epochs = 100
  ))
)
results_h <- evaluate(scheme, algorithms_h, type = "ratings", progress = FALSE)

get_MAE <- function(res){
  map(res, ~.x[3])
}

res_getter_h <- function(results_h){
  n <- names(results_h)
  reslist <- lapply(results_h, getResults)
  v <- map(reslist, get_MAE) %>% unlist()
  
  tibble(
    alg = names(v),
    val = v
  ) %>% 
    mutate(
      fold = str_extract(alg, "\\d{1,2}(?!.*\\d{1,2})"),
      alg = str_remove(alg, fold)
      )
    
  # lapply(reslist, ~tibble(.x[1,]))
}           

res_getter(results_h) %>% 
  ggplot(aes(x=alg, y = val)) +
  geom_boxplot() +
  coord_flip() +
  labs(y = "MAE")
```

I tried other tuning options but nothing performed better than the normal one. More extreme learning rates or regularization led to no results and there seems to be no gains from minor changes. 




Creating Target Profiles
```{r target profile data preprocessing}

artwork_legend <- artworks%>%
  select(Artwork_ID, Title, Style)

target_ave <- left_join(average_ratings_by_piece, artwork_legend)

full_target_df <- left_join(qualtrics_long, target_ave, by = join_by(Artwork_ID))

full_target_df <- full_target_df%>%
  select(participantID, Artwork_ID, Rating, Title, Average_Rating, Style)

full_target_df <- full_target_df %>%
  mutate(Style = case_when(
    Style %in% c("Baroque", "Rococo", "Romanticism") ~ "Baroque",
    Style %in% c("Constructivism", "Art Informel") ~ "Art Informel",
    Style %in% c("Art Nouveau", "Expressionism") ~ "Expressionism",
    Style %in% c("Pop Art", "Op Art") ~ "Pop Art",
    Style %in% c("Abstract Art", "Surrealism", "Cubism") ~ "Abstract",
    Style %in% c("Spatialism", "Minimalism") ~ "Minimalism",
    TRUE ~ Style
  ))
```

```{r}
  unique(full_target_df$Style)
```

```{r eliminating cases with lots of 0 and 100 ratings}

rating_proportions <- full_target_df %>%
  filter(!is.na(Rating)) %>%
  group_by(participantID) %>%
  summarise(
    Total_Ratings = n(),
    Extreme_Ratings = sum(Rating == 0 | Rating == 100, na.rm = TRUE),
    Proportion_Extreme = Extreme_Ratings / Total_Ratings
  )

exclude_participants <- rating_proportions %>%
  filter(Proportion_Extreme > 0.50) %>%
  pull(participantID)

filtered_target_df <- full_target_df %>%
  filter(!participantID %in% exclude_participants)

print(filtered_target_df)

```

```{r calc ecentricity score}


full_target_df <- full_target_df %>%
  mutate(
    Rating = as.numeric(Rating),
    Average_Rating = as.numeric(Average_Rating)
  )

full_target_df1 <- full_target_df %>%
  group_by(Artwork_ID) %>%
  mutate(SD_Average = sd(Rating, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(SD_From_Average = ifelse(!is.na(Rating), abs((Rating - Average_Rating) / SD_Average), NA))


full_target_df2 <- full_target_df1 %>%
  group_by(Artwork_ID) %>%
  mutate(SD_Average = sd(Rating, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(SD_From_Average = ifelse(!is.na(Rating), abs((Rating - Average_Rating) / SD_Average), NA))


eccentricity_df <- full_target_df2 %>%
  group_by(participantID) %>%
  summarise(
    Total_Ratings = n(),
    High_SD_Ratings = sum(!is.na(SD_From_Average) & SD_From_Average >= 1.5)
  ) %>%
  mutate(Eccentricity_Score = High_SD_Ratings / Total_Ratings)


full_target_df2 <- full_target_df2 %>%
  left_join(eccentricity_df, by = "participantID")


print(full_target_df2)


```

```{r descriptives of eccentricity score}

eccentric_des <- full_target_df2 %>%
  summarise(
    Mean = mean(Eccentricity_Score, na.rm = TRUE),
    Median = median(Eccentricity_Score, na.rm = TRUE),
    SD = sd(Eccentricity_Score, na.rm = TRUE),
    Min = min(Eccentricity_Score, na.rm = TRUE),
    Max = max(Eccentricity_Score, na.rm = TRUE),
    Q1 = quantile(Eccentricity_Score, 0.25, na.rm = TRUE),
    Q3 = quantile(Eccentricity_Score, 0.75, na.rm = TRUE)
  )

print(eccentric_des)

ggplot(full_target_df2, aes(x = participantID, y = Eccentricity_Score)) +
  geom_point() +
  labs(title = "Scatter Plot of ParticipantID by Eccentricity Score",
       x = "Participant ID",
       y = "Eccentricity Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```
```{r picking profiles}

full_target_df2 <- full_target_df2 %>%
  mutate(Eccentricity_Category = case_when(
    Eccentricity_Score >= 0 & Eccentricity_Score < 0.025 ~ "0 - 0.025",
    Eccentricity_Score >= 0.025 & Eccentricity_Score < 0.05 ~ "0.025 - 0.05",
    Eccentricity_Score >= 0.05 & Eccentricity_Score < 0.1 ~ "0.05 - 0.1",
    Eccentricity_Score >= 0.1 & Eccentricity_Score < 0.15 ~ "0.1 - 0.15",
    Eccentricity_Score >= 0.15 ~ "0.15+",
    TRUE ~ "NA"
  ))

category_counts <- full_target_df2 %>%
  group_by(Eccentricity_Category) %>%
  summarise(Participant_Count = n_distinct(participantID))


print(category_counts)

set.seed(4)
random_participants <- full_target_df2 %>%
  group_by(Eccentricity_Category) %>%
  sample_n(1, weight = Eccentricity_Score) %>% 
  select(participantID, Eccentricity_Category, Eccentricity_Score)


print(random_participants)
```

```{r}
target_profiles <- full_target_df2%>%
  filter(participantID %in% c(278, 160, 195, 63, 10))


set.seed(1)
target_paintings <- target_profiles %>%
  filter(!is.na(Rating) & !is.na(Average_Rating)) %>%
  group_by(participantID) %>%
  ungroup()


print(target_paintings)
```
```{r test participant 10}
target10<- target_paintings%>%
  filter(participantID == 10)

set.seed(123)
sampled_10 <- target10 %>%
  group_by(Style) %>%
  sample_n(size = 2, replace = FALSE) %>%
  ungroup()
```

```{r test participant 160}
target160<- target_paintings%>%
  filter(participantID == 160)

set.seed(123)
sampled_160 <- target160 %>%
  group_by(Style) %>%
  sample_n(size = 2, replace = FALSE) %>%
  ungroup()
```


```{r test participant 195}
target195<- target_paintings%>%
  filter(participantID == 195)

set.seed(123)
sampled_195 <- target195 %>%
  group_by(Style) %>%
  sample_n(size = 2, replace = FALSE) %>%
  ungroup()
```


```{r test participant 270}
library(dplyr)

# Function to check if a participant can fulfill the conditions
can_fulfill_conditions <- function(data, participant_id) {
  filtered_data <- data %>%
    filter(participantID == participant_id & Eccentricity_Category >= 0.05 & Eccentricity_Category <= 0.1)
  
  if (n_distinct(filtered_data$Style) < length(unique(data$Style))) {
    return(FALSE)
  }
  
  sufficient_rows <- filtered_data %>%
    group_by(Style) %>%
    summarise(count = n()) %>%
    filter(count >= 2)
  
  return(nrow(sufficient_rows) == length(unique(data$Style)))
}

# Function to sample rows for a valid participant
sample_rows_for_valid_participant <- function(data) {
  valid_participant <- NULL
  
  for (participant in unique(data$participantID)) {
    if (can_fulfill_conditions(data, participant)) {
      valid_participant <- participant
      break
    }
  }
  
  if (is.null(valid_participant)) {
    stop("No participant can fulfill the conditions.")
  }
  
  set.seed(123)
  sampled_data <- data %>%
    filter(participantID == valid_participant & Eccentricity_Category == "0.05 - 0.1") %>%
    group_by(Style) %>%
    sample_n(size = 2, replace = FALSE) %>%
    ungroup()
  
  return(sampled_data)
}

# Apply the function to your data
sampled_data <- sample_rows_for_valid_participant(target_paintings)
print(sampled_data)

```

```{r test participant 270}
target270<- target_paintings%>%
  filter(participantID == 270)

set.seed(123)
sampled_270 <- target270 %>%
  group_by(Style) %>%
  sample_n(size = 3, replace = TRUE) %>%
  ungroup()

sampled_270 <- sampled_270 %>%
  distinct(Title, .keep_all = TRUE)

print(sampled_270)
```


```{r test participant 278}
target278<- target_paintings%>%
  filter(participantID == 278)

set.seed(123)
sampled_278 <- target278 %>%
  group_by(Style) %>%
  sample_n(size = 2, replace = FALSE) %>%
  ungroup()
```